---
title: "Mid-course Assignment"
author: "Elizabeth Tullo"
date: '2022-04-10'
---

Here I am loading in all packages I will be using for this assignment.
```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
```


## Part 1: Reading in a data frame

Here I read in my two comma-separated data frames that I will be working with. Both are already in long format (aka tidy) format based on cactus type so no manipulation is needed there.

```{r}
cactus_xylem_data <- read.csv("~/datavisualization/working with ivanna data/cactus_xylem_data.csv")
cross_section_data <- read.csv("~/datavisualization/working with ivanna data/cross_section_data.csv")
```




## Part 2: Generate Summary Statistics

Here I generate summaries of the three different data frames. First I generate a summary for cactus_xylem_data, then cross_section_data.

```{r}
summary(cactus_xylem_data)
summary(cross_section_data)
```






## Part 3: Manipulate the Data

### Manipulation 1: Adding Columns

Here I calculate the vessel_radius for the vessels in Cactus_xylem_data based on the column "Vessel_perimeter".
```{r}
vessel_radius <- (cactus_xylem_data$Vessel_perimeter) / (2 * pi)
```


Here I add the Vessel_radius as a column to the Cactus_xylem_data dataset.
```{r}
cactus_xylem_data$vessel_radius <- vessel_radius
```


Here I calculate the summary statistics for this new column.
```{r}
summary(cactus_xylem_data$vessel_radius)
```



**I then repeat the above steps to calculate the cross_section_radius for the Cross_section_data**

Here I calculate the Cross_section_radius for the cross_sections in cross_section_data based on the column "Cross_section_perimeter".
```{r}
cross_section_radius <- (cross_section_data$Cross_section_perimeter) / (2 * pi)
```


Here I add the cross_section_radius as a column to the Cross_section_data dataset.
```{r}
cross_section_data$cross_section_radius <- cross_section_radius
```


Here I calculate the summary statistics for this new column.
```{r}
summary(cross_section_data$cross_section_radius)
```




### Manipulation 2: Merging the two data sets

Here I full-join the two data frames cactus_xylem_data and cross_section_data into cactus_xylem_and_cross_section based on "Cactus" type and "Height_from_apex".

```{r}
cactus_xylem_and_cross_section <- full_join(cactus_xylem_data, cross_section_data, by = c("Cactus", "Height_from_apex"))
```


Here I calculate the summary statistics for the new merged data set cactus_xylem_and_cross_section.
```{r}
summary(cactus_xylem_and_cross_section)
```


## Part 4: Generate ggplots


### Geom 1: Histogram


### Geom 2: Scatter Plot


### Geom 3: Jitter
